\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{A Factorial Design for Separating Persona and Temperature Effects in
LLM-Based Fuzzy Emotion Scoring}

\author{
\IEEEauthorblockN{Naruki Shirahama}
\IEEEauthorblockA{\textit{Shimonoseki City University} \\
Shimonoseki, Yamaguchi, Japan \\
nshirahama@ieee.org}
\and
\IEEEauthorblockN{Yuma Yoshimoto}
\IEEEauthorblockA{\textit{NIT (KOSEN), Kitakyushu College} \\
Kitakyushu, Fukuoka, Japan \\
yoshimoto@kct.ac.jp}
\and
\IEEEauthorblockN{Naofumi Nakaya}
\IEEEauthorblockA{\textit{Juntendo University} \\
Urayasu, Chiba, Japan \\
n.nakaya.ac@juntendo.ac.jp}
\and
\IEEEauthorblockN{Junji Nishino}
\IEEEauthorblockA{\textit{Tamagawa University} \\
Machida, Tokyo, Japan \\
nishinojunji@eng.tamagawa.ac.jp}
\and
\IEEEauthorblockN{Satoshi Watanabe}
\IEEEauthorblockA{\textit{Shizuoka Institute of Science and Technology} \\
Fukuroi, Shizuoka, Japan \\
satoshi\_watanabe@ieee.org}
}

\maketitle

\begin{abstract}
Large language models are increasingly used as scalable judges of literary
emotion, but persona prompts and decoding temperature are often varied together.
This coupling aliases interpretive stance with stochastic decoding control. We
introduce a fully crossed persona-temperature design for Japanese literary
emotion scoring and evaluate score, validity, and normalized fuzzy entropy
within the same balanced factorial structure. The main run uses six
contemporary models, three literary texts, four persona stances, four
temperature levels, and five repeats per cell, yielding 1440 responses and 5760
emotion-level scores. After retrying only four initially invalid rows, all
records passed the structured-output gate. Within each model-text-emotion unit,
balanced decomposition summaries show that mean scores are largely
persona-driven and nearly additive, whereas normalized fuzzy entropy carries
larger persona-temperature interaction burdens, especially for anger and
surprise. The results show that independent crossing is needed not only for
cleaner score attribution, but also for detecting when membership-overlap
entropy remains jointly sensitive to prompting and sampling controls.
\end{abstract}

\begin{IEEEkeywords}
large language models, persona prompting, temperature, fuzzy entropy, emotion scoring, factorial design
\end{IEEEkeywords}

\section{Introduction}
LLM-based emotion scoring offers a practical way to compare affective
interpretations across models and prompt settings. In prior work on emotional
individuality, persona prompting and decoding temperature were used to induce
diversity in model responses. However, when a specific persona is paired with a
specific temperature, the resulting score variation cannot be cleanly assigned
to either interpretive stance or decoding stochasticity. This is a design
limitation rather than only a statistical limitation: no downstream model can
separate two factors that were not independently varied.

This aliasing problem is not specific to literary emotion scoring. Many LLM
evaluation protocols combine role prompts, response-style instructions,
sampling parameters, and repeated generations. If these controls are changed
together, later analysis may attribute an observed score difference to the
wrong control. The methodological issue is therefore to separate prompt-side
interpretive controls from decoding-side stochastic controls before fitting or
summarizing downstream outcomes.

This paper addresses that limitation with a fully crossed persona-temperature
design, following the standard logic of factorial designs and interactions
\cite{montgomery2019design}. We hold the text set and response schema fixed,
express persona only as an interpretive stance in the prompt, and pass
temperature only as an API configuration parameter. The central question is
whether persona effects, temperature effects, and their interaction can be
described separately for score-level emotion ratings and fuzzy entropy derived
from those ratings.

Our contribution is methodological and empirical. First, we define a
conference-scale factorial protocol for LLM-based literary emotion scoring.
Second, we compute balanced effect components for persona, temperature, and
persona-temperature interaction under the same cell structure for score and
entropy. Third, we update the fuzzy entropy pipeline to use a smooth
sigmoid-composed membership family as the primary specification, while retaining
the prior piecewise-linear family as a sensitivity baseline. The intended claim
is not that one membership function is uniquely correct, but that the factorial
comparison is made under an explicit and reproducible membership specification.

\section{Related Work}
LLM emotion benchmarks have evaluated affective understanding, empathy, and
emotion recognition from several perspectives \cite{sabour2024emobench,
paech2024eqbench}. These efforts establish that LLMs can be compared on
emotion-related tasks, but they do not by themselves resolve how prompt persona
and sampling parameters affect continuous score formation. Our setting is
narrower: we focus on repeated literary emotion scoring under a controlled
factorial design.

Persona prompting and decoding temperature are scientifically different
controls. Persona instructions are prompt-side role or stance manipulations,
while temperature is a generation-side sampling parameter. Surveys and empirical
studies of LLM personas show that assigned personas can systematically affect
model behavior \cite{tseng2024personaSurvey,jiang2024personallm}. Work on
stochastic decoding and temperature shows that sampling controls affect output
diversity and task performance \cite{holtzman2020curious,renze2024temperature}.
When these controls are paired rather than independently crossed, the resulting
variation is design-aliased.

Fuzzy-set theory provides a natural language for representing degrees of
membership rather than crisp class labels \cite{zadeh1965fuzzy}. Shannon's
entropy form \cite{shannon1948} and nonprobabilistic fuzzy entropy
\cite{deluca1972nonprobabilistic} motivate entropy-like measures of fuzziness,
although membership values here are not treated as probability distributions.
Prior EQ-Fuzzy work used Low, Medium, and High membership values over 0--100
emotion intensity scores to summarize score placement in LLM scoring
\cite{shirahama2026mdpi}. Because literary texts are a meaningful domain for
sentiment and emotion analysis \cite{kim2019literaryEmotion}, we use Japanese
literary scoring as a fixed task panel. In this paper, entropy is interpreted as
membership-overlap entropy under an explicit Low/Medium/High membership family,
not as a universal measure of human emotional ambiguity.

\section{Method}
Let $x_{m,t,p,\tau,r,e}$ denote a 0--100 score produced by model $m$ for text
$t$, persona $p$, temperature $\tau$, repeat $r$, and emotion $e$. The primary
factorial cell is $(m,t,p,\tau)$, with four emotion dimensions recorded for each
response. For each model-text-emotion unit, we summarize the $4 \times 4$
persona-temperature mean table. Let $\bar{y}_{p\tau}$ denote the repeat mean in
persona-temperature cell $(p,\tau)$, with marginal means
$\bar{y}_{p\cdot}$ and $\bar{y}_{\cdot\tau}$ and grand mean
$\bar{y}_{\cdot\cdot}$. The balanced components are:
\begin{equation}
SS_{\mathrm{persona}} =
4\sum_p(\bar{y}_{p\cdot}-\bar{y}_{\cdot\cdot})^2,
\end{equation}
\begin{equation}
SS_{\mathrm{temp}} =
4\sum_\tau(\bar{y}_{\cdot\tau}-\bar{y}_{\cdot\cdot})^2,
\end{equation}
\begin{equation}
SS_{\mathrm{persona}\times\mathrm{temp}} =
\sum_{p,\tau}
(\bar{y}_{p\tau}-\bar{y}_{p\cdot}-\bar{y}_{\cdot\tau}
+\bar{y}_{\cdot\cdot})^2.
\end{equation}
We report the relative interaction burden
\begin{equation}
B =
\frac{SS_{\mathrm{persona}\times\mathrm{temp}}}
{SS_{\mathrm{persona}} + SS_{\mathrm{temp}} +
SS_{\mathrm{persona}\times\mathrm{temp}}},
\end{equation}
and separability share $1-B$. Low interaction burden indicates that persona and
temperature are practically separable for the outcome under analysis. The
decomposition is descriptive: it quantifies balanced factorial variation in the
observed panel, not causal effects of API temperature across providers.

\subsection{Design Rationale}
The model and text panels are treated as selected fixed panels. This choice
matches the conference-paper goal: to test whether a deconfounded protocol can
estimate separability within a concrete contemporary model panel, rather than
to infer population-level effects over all possible models or literary texts.
Temperature is treated as a categorical factor because API temperature scales
are provider-mediated and need not have identical behavioral meaning across
routes. The five repeats per cell provide a conference-scale stability check;
larger repeat depths are reserved for journal extension. Finally, the response
schema and manifest rows are immutable analysis units. Failed rows are repaired
only by targeted retry of the same manifest rows, so the repair policy does not
change valid observations or alter the factorial design.

For fuzzy entropy, each score is mapped to Low, Medium, and High membership
values. The primary family, denoted \texttt{sigmoid\_s\_v1}, preserves the
prior semantic breakpoints: Low decreases over 0--30, Medium rises over 10--30,
has a plateau over 30--50, and falls over 50--70, and High rises over 50--100.
Unlike the prior piecewise-linear family \cite{shirahama2026mdpi}, each
transition interval $[\ell,h]$ is first mapped by a sigmoid with midpoint
$(\ell+h)/2$ and gain $2\log 9/(h-\ell)$, and then passed through a smooth
S-function. This sigmoid-composed S-shape follows recent membership-function
design considerations that prefer smooth transitions while retaining bounded
endpoint behavior \cite{umano2025membership}. We compute raw fuzzy entropy
$H_{\mathrm{raw}} = -\sum_k \mu_k \log_2 \mu_k$ with $0\log 0=0$, and report
the normalized value $H^* = H_{\mathrm{raw}}/H_{\max}$, where $H_{\max}$ is
deterministically estimated for each membership family on a dense score grid.
Absolute entropy magnitudes are therefore interpreted only within a fixed
membership specification; the legacy family is retained as a sensitivity
baseline.
\input{../../artifacts/scis2026/main_figures_v1/figure1a_prior_diagonal_design}
\input{../../artifacts/scis2026/main_figures_v1/figure1b_factorial_design}

\section{Experiments}
The main experiment used six contemporary LLMs, three Japanese literary texts,
four persona stances, four temperature levels ($0.1$, $0.4$, $0.7$, and $0.9$),
and five repeats per condition. This gives
$6 \times 3 \times 4 \times 4 \times 5 = 1440$ response trials. Each response
was required to return four emotion scores and corresponding reasons in the
fixed response schema. The model panel consisted of GPT-5.4, Claude Sonnet 4.5,
Gemini 2.5 Pro, Grok 4.20, DeepSeek V3.2, and Llama 4 Maverick. Temperature was
never mentioned in the prompt text; persona was expressed only as the
interpretive stance.

\begin{table}[t]
\caption{SCIS main-run experimental panel.}
\label{tab:scis-design-panel}
\centering
\begin{tabular}{ll}
\hline
Component & Levels \\
\hline
Models & 6 provider-accessed LLMs \\
Texts & 3 Japanese literary texts \\
Personas & freshman, scholar, poet, robot \\
Temperatures & 0.1, 0.4, 0.7, 0.9 \\
Repeats & 5 per cell \\
Emotions & interest, surprise, sadness, anger \\
Trials & 1440 responses / 5760 scores \\
\hline
\end{tabular}
\end{table}

The three texts were ``The Pocket Watch'' and ``The Money and the Pistol'' by
Kyusaku Yumeno, and ``The Tattered Ostrich'' by Kotaro Takamura. The four
persona stances were a college freshman, a literary scholar, an emotive poet,
and an emotionless robot. Models were accessed through provider routes recorded
in the manifest; display names are used here for readability, while exact model
identifiers, prompt versions, parser versions, and manifest hashes are retained
in the reproducibility artifacts.

The initial main run contained four invalid rows, corresponding to an initial
failure rate of $4/1440=0.28\%$. We repaired these by retrying only the failed
manifest rows and merging the successful retry records into a separate repaired
raw file, leaving the original raw output unchanged. No valid row was rerun or
selectively filtered. The repaired run reached complete structured-output
validity, as shown in Table~\ref{tab:scis-validity}.
\input{../../artifacts/scis2026/main_tables_v1/table1_run_validity}

\section{Results}
Table~\ref{tab:scis-bootstrap-ci} summarizes the primary decomposition with
within-cell bootstrap intervals. The resampling keeps model, text, persona, and
temperature cells fixed and resamples the five repeats within each cell,
preserving the four emotion scores from the same response as a block. These
intervals are conditional on the selected models, texts, personas, and
temperatures; they quantify repeat-level stability within the realized panel,
not uncertainty over the population of possible models or literary texts. For
mean score, persona accounts for most of the decomposed variation across all
four emotions. The mean persona share is above 90\% for interest, sadness, and
surprise, and remains above 80\% for anger. Temperature shares are comparatively
small. Score-level interaction burdens range from 3.4\% for sadness to 13.2\%
for anger, so score variation is largely separable within the crossed subtables.
\input{../../artifacts/scis2026/main_bootstrap_v1/effect_summary_bootstrap_ci}

The entropy results show a different pattern. Although persona remains the
largest average component for $H^*$, interaction burdens are higher than for
score in every emotion. They reach 28.3\% for anger and 19.8\% for surprise,
compared with 13.2\% and 7.0\% for the corresponding score summaries. This
means that score means and entropy descriptors answer different questions:
scores identify the dominant interpretive stance effect, while entropy reveals
where the placement of scores relative to Low, Medium, and High remains jointly
sensitive to persona and temperature.

This score-entropy contrast is the main empirical value of the factorial
design. If only mean scores were reported, the conclusion would be that persona
and temperature are mostly separable in the selected panel. The entropy analysis
shows where that separability weakens: the same score table can remain nearly
additive while the fuzziness of score placement retains a larger interaction
component. A diagonal design that observes only one temperature per persona
would not expose this metric-specific difference.
\input{../../artifacts/scis2026/main_figures_v1/figure2a_entropy_heatmap}
\input{../../artifacts/scis2026/main_figures_v1/figure2b_score_heatmap}

Fig.~\ref{fig:scis-model-metric-heatmap} summarizes the same contrast at the
model level. The score column is generally lighter than the entropy column,
showing that the larger interaction burden for $H^*$ is not limited to a single
provider. DeepSeek V3.2 and Gemini 2.5 Pro show comparatively high mean
interaction burden for normalized fuzzy entropy, while Claude Sonnet 4.5 and
Grok 4.20 show low interaction burden for score. The detailed model-level
shares are retained in the reproducibility artifacts.
\input{../../artifacts/scis2026/main_figures_v1/figure3_model_metric_interaction_heatmap}

The entropy sensitivity check in Table~\ref{tab:scis-entropy-sensitivity}
compares the primary smooth membership family with the prior piecewise-linear
baseline. The cell-mean correlation is high, indicating that the updated
membership family does not create an unrelated entropy signal. At the same
time, the nonzero absolute differences confirm that membership-family choice
affects entropy magnitude. We therefore interpret $H^*$ comparatively under a
fixed family, and use the legacy family as a sensitivity baseline rather than as
a competing primary endpoint.
\input{../../artifacts/scis2026/main_tables_v1/table4_entropy_sensitivity}

Additional reviewer-risk diagnostics support the same interpretation. The four
original failed rows were all repaired by targeted retry, and the repaired file
has zero failed rows. Leave-one-model-out analysis changed mean interaction
burden by at most 0.034, and the cell-mean correlation between the primary
smooth entropy family and the legacy linear family was 0.933. These checks do
not prove generality, but they reduce the risk that the central score-entropy
contrast is an artifact of one provider or one membership implementation.

Detailed point-estimate tables and top-interaction case lists are retained in
the reproducibility artifacts for appendix selection. The main paper instead
uses the CI table and representative heatmaps to avoid duplicating the same
decomposition in multiple large tables.

\section{Discussion}
The experiment is best understood as a design diagnostic rather than a model
leaderboard. The six models are not compared to decide which model is most
emotionally accurate; they form a contemporary panel for testing whether prompt
stance and decoding temperature can be separated in a structured scoring
pipeline. This framing matters because the same model can look stable under one
metric and interaction-sensitive under another. The factorial protocol exposes
that distinction by forcing every persona to be observed at every temperature.

The main substantive implication is metric-specific separability. Mean scores
are mostly additive in persona and temperature within the selected
model-text-emotion subtables. This supports the practical use of persona as a
strong prompt-side control for score-level interpretation. Normalized fuzzy
entropy, however, retains larger interaction burdens. Thus entropy should not
be treated as a redundant transform of the score mean. It captures how scores
move relative to Low, Medium, and High membership regions, and that placement
can remain jointly sensitive to prompt framing and sampling even when score
means are nearly separable.

This distinction also explains why the membership specification is part of the
method rather than an incidental implementation detail. The paper does not claim
that the sigmoid-S family is the uniquely correct representation of emotion
intensity. Instead, it fixes a smooth and reproducible Low/Medium/High
specification, reports normalized entropy within that family, and checks the
legacy linear family as a sensitivity baseline. Under this interpretation,
fuzzy entropy is a controlled descriptor of membership overlap in the scoring
pipeline, not a direct measurement of human emotional uncertainty.

\section{Limitations and Extension}
The results are design-conditional. They describe the selected six-model panel,
three Japanese literary texts, four persona stances, four temperature levels,
parser, response schema, and primary membership family. They do not establish
human validity, cross-lingual generality, or provider-independent causal effects
of temperature. The selected texts are meaningful literary emotion stimuli, but
they are not a random sample of Japanese literature. Likewise, display model
names should not be read as stable scientific objects independent of provider
updates.

The repeat depth is intentionally modest. With $r=5$, the bootstrap intervals
summarize repeat-level stability within the realized panel; they do not
quantify uncertainty over future model versions or broader text collections.
The fuzzy entropy analysis is also specification-dependent. We address this by
normalizing entropy within the chosen membership family and retaining the
legacy linear family as a sensitivity baseline, but no membership family should
be interpreted as uniquely correct. A journal extension should increase repeat
depth, broaden the model and text panels, add human-reference validation, and
test robustness under additional membership and model-update conditions.

\section{Conclusion}
We presented a fully crossed persona-temperature design for LLM-based fuzzy
emotion scoring and applied it to a six-model Japanese literary emotion
experiment. The main result is that score variation is largely separable and
persona-driven within the selected model-text-emotion subtables, whereas
normalized fuzzy entropy retains stronger
persona-temperature interaction, particularly for anger and surprise. This
distinction would be obscured in a diagonal design that bundles persona and
temperature together. The practical lesson is that prompt-side interpretive
controls and decoding-side stochastic controls should be independently varied
when they are both used in LLM-based affective scoring. Under such a design,
mean scores and membership-overlap entropy can reveal different separability
profiles even when they are computed from the same structured responses.

\section*{Acknowledgment}
This work was supported in part by the EQ-Fuzzy research line under KAKENHI.

\begin{thebibliography}{00}

\bibitem{zadeh1965fuzzy}
L. A. Zadeh, ``Fuzzy sets,'' \emph{Information and Control}, vol. 8, no. 3,
pp. 338--353, 1965, doi: 10.1016/S0019-9958(65)90241-X.

\bibitem{shannon1948}
C. E. Shannon, ``A mathematical theory of communication,'' \emph{Bell System
Technical Journal}, vol. 27, no. 3, pp. 379--423, 1948, doi:
10.1002/j.1538-7305.1948.tb01338.x.

\bibitem{deluca1972nonprobabilistic}
A. De Luca and S. Termini, ``A definition of a nonprobabilistic entropy in the
setting of fuzzy sets theory,'' \emph{Information and Control}, vol. 20, no. 4,
pp. 301--312, 1972, doi: 10.1016/S0019-9958(72)90199-4.

\bibitem{shirahama2026mdpi}
N. Shirahama, Y. Yoshimoto, N. Nakaya, and S. Watanabe, ``Mathematical
Framework for Characterizing Emotional Individuality in Large Language Models:
Temperature Control, Fuzzy Entropy, and Persona-Based Diversity Analysis,''
\emph{Mathematics}, vol. 14, no. 7, Art. no. 1224, 2026, doi:
10.3390/math14071224.

\bibitem{montgomery2019design}
D. C. Montgomery, \emph{Design and Analysis of Experiments}, 10th ed. Hoboken,
NJ, USA: Wiley, 2019.

\bibitem{tseng2024personaSurvey}
Y.-M. Tseng, Y.-C. Huang, T.-Y. Hsiao, W.-L. Chen, C.-W. Huang, Y. Meng, and
Y.-N. Chen, ``Two tales of persona in LLMs: A survey of role-playing and
personalization,'' in \emph{Findings of ACL: EMNLP 2024}, 2024,
pp. 16612--16631, doi: 10.18653/v1/2024.findings-emnlp.969.

\bibitem{jiang2024personallm}
H. Jiang, X. Zhang, X. Cao, C. Breazeal, D. Roy, and J. Kabbara,
``PersonaLLM: Investigating the ability of large language models to express
personality traits,'' in \emph{Findings of ACL: NAACL 2024}, 2024,
pp. 3605--3627, doi: 10.18653/v1/2024.findings-naacl.229.

\bibitem{holtzman2020curious}
A. Holtzman, J. Buys, L. Du, M. Forbes, and Y. Choi, ``The curious case of
neural text degeneration,'' in \emph{Proc. ICLR}, 2020.

\bibitem{renze2024temperature}
M. Renze, ``The effect of sampling temperature on problem solving in large
language models,'' in \emph{Findings of ACL: EMNLP 2024}, 2024,
pp. 7346--7356, doi: 10.18653/v1/2024.findings-emnlp.432.

\bibitem{umano2025membership}
M. Umano, ``Some considerations on membership functions of fuzzy sets,'' in
\emph{Proc. 41st Fuzzy System Symposium}, 2025, pp. 70--74, doi:
10.14864/fss.41.0\_70.

\bibitem{sabour2024emobench}
S. Sabour \emph{et al.}, ``EmoBench: Evaluating the emotional intelligence of
large language models,'' in \emph{Proc. ACL 2024}, 2024, pp. 5986--6004,
doi: 10.18653/v1/2024.acl-long.326.

\bibitem{paech2024eqbench}
S. J. Paech, ``EQ-Bench: An emotional intelligence benchmark for large language
models,'' arXiv:2312.06281, 2023.

\bibitem{kim2019literaryEmotion}
E. Kim and R. Klinger, ``A survey on sentiment and emotion analysis for
computational literary studies,'' \emph{Zeitschrift f\"ur digitale
Geisteswissenschaften}, no. 4, pp. 1--23, 2019, doi:
10.17175/2019\_008\_v2.

\end{thebibliography}

\end{document}
